---
title: "Descriptive statistics and Graphical Analysis FSt.02"
author: "Lloyd Adu-Kumi"
date: "2025-11-14"
categories: [Food Statistics, Analysis]
image: true
---

## Descriptive Statistics

Statistics for Food Scientists

***Chapter Commentary:***

Maria is given a tour of the BBQ sauce production plant to understand how manufacturing works (Understanding the food production flow makes the data more relatable!😉). The plant manager, Lisa, takes her from ingredient handling to mixing, heating, filling, and hourly viscosity checks for quality control. Lisa introduces a new high-speed filling line and asks Maria to produce data to confirm whether the new system truly improves performance.

To describe our data, we utilise summary statistics to give an idea of viscosity levels in the individual production line. I love R and Python, but let's try R for this book😊!

Let's first import our mock data

```{r}
#| label: Load Data
#| message: false
#load packages
pacman::p_load(
  rio,  #for importing data
  here, #for relative file paths
  skimr, #for reviewing the data
  lubridate, # for date cleaning
  janitor, #for cleaning data
  flextable, #for making pretty tables
  gtsummary, # creating tables
  scales, #percents in tables
  readxl,
  moments,
  tidyverse #for data management and viz
  
)

#Import Data
bbq <- import("wwm.xlsx") %>% 
  clean_names() %>%   #to standardize the names of the columns
  mutate(
    across(
      starts_with("date"),
      ~ mdy_hm(.)
    )
  )
head(bbq) #Let's return the top 6 rows
 
```

Maybe we could pivot_longer to consolidate the data points

```{r}
bbq <- bbq %>% 
  pivot_longer(
    cols = starts_with("line"),
    names_to = "line",
    values_to = "viscosity"
  ) %>% 
  pivot_longer(
    cols = starts_with("date"),
    names_to = NULL,
    values_to = "date"
  ) 

head(bbq)
```

Let's visualize the data into histograms

```{r}
#| label: Histogram Chunk
#| message: false
#| warning: false
ggplot(data = bbq,
       mapping = aes(
         x = viscosity
       ))+
  geom_histogram(bins = 10, fill = "lightblue")+
  labs(
    y = "Frequency"
  )+
  facet_wrap(~line)+
  theme_minimal()

```

```{r}
bbq %>% 
  group_by(line) %>% 
  summarise(
   Mean = mean(viscosity),
    Median = median(viscosity),
    `Standard deviation` = sd(viscosity),
    Kurtosis = kurtosis(viscosity),
    Skewness = skewness(viscosity),
    Range = max(viscosity) - min(viscosity),
    Minimum = min(viscosity),
    Maximum = max(viscosity)
  ) %>% 
  qflextable()
```
